The quench action approach in finite integrable spin chains

### 2015-12-05 QAMC_draft/fs_qa_v6.tex
05/12/2015

\title{The quench action approach for integrable models: A Monte Carlo study}

\author{Authors}
%\address{$^1$ International School for Advanced Studies (SISSA),
%Via Bonomea 265, 34136, Trieste, Italy,
%INFN, Sezione di Trieste}

@@

fdasfa
\end{abstract}

@@

Understanding the out-of-equilibrium dynamics in {\it isolated} quantum 
many-body systems is one of the most intriguing research topics in 
contemporary physics, both experimentally~\cite{bloch-2008,greiner-2002,
kinoshita-2006,hofferberth-2007,trotzky-2012,gring-2012,cheneau-2012,
schneider-2012,kunhert-2013,langen-2013,meinert-2013,fukuhara-2013,
ronzheimer-2013,braun-2014,langen-2015} and theoretically~\cite{polkovnikov-2011}. 
The most investigated protocol is that of the {\it quantum quench}, in 
which a system is initially prepared in an eigenstate $|\Psi_0\rangle$ of 
a many-body hamiltonian ${\mathcal H}_0$. Then a global parameter is 
suddenly changed, and the system is let to evolve unitarily under a new 
hamiltonian ${\mathcal H}$. There is now compelling evidence that at long 
times after the quench an equilibrium steady state arises (see for instance 
Ref.~\cite{kinoshita-2006}), although its nature is not fully understood yet. 
In the thermodynamic limit, due to dephasing, it is natural to expect that 
the equilibrium value of any local observable ${\mathcal O}$ is described 
by the so-called diagonal ensemble as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle_{DE}=\sum\limits_{\lambda}|\langle\Psi_0|\lambda
\rangle|^2\langle\lambda|{\mathcal O}|\lambda\rangle,
\end{equation}
%
where the sum is over the eigenstates $|\lambda\rangle$ of the post-quench 
hamiltonian ${\mathcal H}$. Moreover, for generic, i.e., non-integrable, models 
the Eigenstate Thermalization Hypothesis~\cite{deutsch-1991,srednicki-1994} 
(ETH) implies that the diagonal ensemble~\eref{d-ensemble} becomes equivalent 
to the usual Gibbs (thermal) ensemble.

In integrable models, however, the presence of an extensive number of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the 
onset of thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,
rigol-2008} that in this situation the post-quench steady-state can be described 
by the so-called Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved quantities, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
Note that for the diagonal ensemble~\eref{d-ensemble}  one has trivially that $\langle
{\mathcal O}\rangle_{DE}=\langle\Psi_0|{\mathcal O}|\Psi_0\rangle$. We anticipate that 
this provides  useful sum rules that will be the main object of interest in the paper. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2011,
calabrese-2012,fagotti-2013,kcc14,kcc14a}, whereas in interacting ones the scenario is 
still not settled. Note that a recent work~\cite{ilievski-2015a} suggests that the $GGE$ 
description of the steady state  is complete provided that the so-called quasi-local 
charges~\cite{prosen-2014,pereira-2014,ilievski-2015} are included in~\eref{gge}.

Valuable insights into this issue have been provided by the so-called quench-action 
approach~\cite{caux-2013}. For Bethe ansatz integrable models this method allows one 
to construct the diagonal ensemble~\eref{d-ensemble} directly in the thermodynamic 
limit, provided that the overlaps $\langle\lambda|\Psi_0\rangle$ are known. The 
physical idea is that for large system sizes it is possible to approximate the sum 
over the eigenstates in~\eref{d-ensemble} using a saddle point argument. Specifically, 
one starts with rewriting~\eref{d-ensemble} as 
%
\begin{equation}
\label{de1}
\langle{\mathcal O}\rangle_{DE}= \sum\limits_{\lambda}\rho^{DE}\langle\lambda|
{\mathcal O}|\lambda\rangle,\quad\textrm{with}\quad\rho^{DE}=\exp(-2\Re{\mathcal 
E}(\lambda)), 
\end{equation}
%
where ${\mathcal E}(\lambda)\equiv-\log\langle\lambda|\Psi_0\rangle$, and $\Re$ 
denotes the real part. For typical initial states $|\Psi_0\rangle$
one has ${\mathcal E}\propto L$, with $L$ the system size. This reflects the 
vanishing of the overlaps as $\langle\Psi_0|\lambda\rangle\propto e^{-c L}$.  
As in the standard Thermodynamic Bethe Ansatz (TBA)~\cite{taka-book}, the 
extensivity of ${\mathcal E}$ suggests that in the thermodynamic limit the 
sum in~\eref{de1} is dominated by a saddle point. Remarkably, for some Bethe 
ansatz solvable models and for simple-enough initial states, it is possible to 
determine this saddle point {\it analytically}. This has been done succesfully, 
for instance, for the quench from the Bose-Einstein condensate~\cite{de-nardis-2014} 
(BEC) in the Lieb-Liniger model and for the quench from the N\'eel 
state~\cite{wouters-2014,brockmann-2014} in the $XXZ$ spin chain. Interestingly, 
it has been suggested recently that the quench action approach could allow to 
reconstruct the full relaxation dynamics to the steady state~\cite{de-nardis-2015a,
de-nardis-2015}.

\paragraph*{Outline of the results.} 
In this paper, by combining exact Bethe ansatz and Monte Carlo techniques, we investigate 
the diagonal ensemble and the quench action approach in {\it finite} size integrable models. 
We focus on the spin-$1/2$ isotropic Heisenberg ($XXX$) chain with $L$ sites, which is 
defined by the Hamiltonian 
%
\begin{equation}
\label{xxx-ham}
{\mathcal H}\equiv J\sum\limits_{i=1}^L\left[\frac{1}{2}(S_i^+S^-_{i+1} 
+S_i^{-}S_{i+1}^+)+S_i^zS_{i+1}^z-\frac{1}{4}\right].
\end{equation}
%
Here $S^{\pm}_i\equiv (\sigma_i^x\pm i\sigma_i^y)/2$ are spin operators acting on the 
site $i$, $S_i^z\equiv\sigma_i^z/2$, and $\sigma^{x,y,z}_i$ the Pauli matrices. We 
fix $J=1$ in~\eref{xxx-ham} and use periodic boundary conditions, identifying sites 
$L+1$ and $1$. The total magnetization $S_{T}^z\equiv\sum_iS_i^z=L/2-M$, with $M$ number 
of down spins (particles), commutes with~\eref{xxx-ham}, and it is used to label its 
eigenstates. The $XXX$ chain is Bethe ansatz solvable~\cite{bethe-1931} and its eigenstates 
are in one-to-one correspondence with the solutions of the so-called Bethe equations 
(see~\ref{sec:1.2}). We restrict ourselves to quenches from the zero-momentum N\'eel 
state $|N\rangle$ and the Majumdar-Ghosh state $|MG\rangle$, which are defined as  
%
\begin{equation}
\label{psi0}
|N\rangle\equiv\Big(\frac{|N_1\rangle
+|N_2\rangle}{\sqrt{2}}\Big), \qquad 
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
Here $|N_1\rangle\equiv |\uparrow\downarrow\rangle^{\otimes L/2}$, and $|N_2\rangle
\equiv |\downarrow\uparrow\rangle^{\otimes L/2}$, with $\otimes$ denoting the tensor 
product. Note that both $|N\rangle$ and $|MG\rangle$ are invariant under one site 
translations. Our study relies on the analytical knowledge of the overlaps between 
the N\'eel and Majumdar-Ghosh state and the eigenstates of~\eref{xxx-ham}~\cite{
brockmann-2014,brockmann-2014a,brockmann-2014b,brockmann-2014c,pozsgay-2014a,
piroli-2014}.

We first present a detailed overview of the distribution of the overlaps between 
the initial states and the eigenstates of the model (overlap distribution function). 
Precisely, we provide numerical results for all the overlaps for finite chains up 
to $L\le 38$. Our results are obtained exploiting the analytical formulas 
presented in Ref.~\cite{brockmann-2014} and~\cite{pozsgay-2014a}. Crucially, 
we restrict ourselves to a truncated Hilbert space, considering only eigenstates 
of~\eref{xxx-ham} that do not contain zero-momentum strings. Physically, 
zero-momentum strings correspond to eigenstates amplitudes containing multi-particle  
bound states which are fully delocalized in the chain. From the Bethe ansatz perspective, 
in the thermodynamic limit the presence of zero-momentum strings leads to fictitious 
singularities in the overlap formulas. Dealing with these singularities is 
a formidable task that requires detailed knowledge of the solutions of the 
Bethe equations, and it can be done only in very simple cases, e.g., for small 
chains. On the other hand, it has been argued that zero-momentum strings are irrelevant 
in the thermodynamic limtit~\cite{brockmann-2014a}.

For a finite chain here we find that for both the N\'eel and the Majumdar-Ghosh 
cases, the fraction of eigenstates that do not contain zero-momentum strings 
is vanishing in the thermodynamic limit, meaning that for finite large-enough 
systems the vast majority of the eigenstates contain zero momentum strings. Specifically, 
the total number of eigenstates without zero-momentum strings is given in terms of the 
chain size $L$ by simple combinatorial formulas that we provide. We investigate the 
effect of the Hilbert space truncation on the diagonal ensemble focusing on the  
sum rules $\langle Q_j\rangle_{DE}=\langle\Psi_0|Q_j|\Psi_0\rangle$ for the conserved 
quantities $Q_j$. We numerically demonstrate that truncating the diagonal ensemble to 
eigenstates with no zero-momentum strings leads to striking violations of the sum 
rules. Precisely, we numerically observe that the truncated diagonal ensemble 
average $\langle Q_j\rangle$ vanish in the thermodynamic limit reflecting the 
vanishing behavior of the fraction of finite-momentum strings eigenstates. 
This allows us to conclude that for finite chaings eigenstates corresponding 
to zero-momentum strings cannot be trivially neglected when considering 
diagonal ensemble averages.

However, we provide numerical evidence that the correct diagonal ensemble 
sum rules can be recovered in the thermodynamic limit by appropriately 
reweighting the contribution of the finite-momentum strings eigenstates.  
We focus only on the situation with the N\'eel state (cf.~\eref{psi0}) as 
initial state. The key idea is to sample the eigenstates with no zero-momentum 
strings with the diagonal ensemble probability measure $\rho^{DE}$ (cf.~\eref{de1}). 
To this purpose we develop a Monte Carlo scheme, which is a generalization of 
the approach presented in Ref.~\cite{alba-2015} to simulate the Generalize 
Gibbs Ensemble. The method is based on the knowledge of the N\'eel overlaps 
$\langle\lambda|N\rangle$ and on the knowledge of the Hilbert space structure 
of the $XXX$ chain in the Bethe ansatz formalism. The approach allowed us to 
simulate effectively chain with $L\lesssim 60$, although larger systems 
sizes can in principle be reached.

Strikingly, although for small chains violations of the conserved quantities 
sum rules are present, these violations vanish in the thermodynamic limit 
and the sum rules are restored. This implies that the only effect of the 
Hilbert space truncation is to introduce scaling corrections. In the quench 
action language, this means that the eigenstates corresponding to no 
zero-momentum strings contain enough physical information about saddle point. 
This is numerically confirmed by extracting the so-called saddle point 
root distributions, which in the Bethe ansatz language fully characterize 
the diagonal ensemble averages in the thermodynamic limit. In the numerical  
 approach these are obtained from the histograms of the Bethe ansatz 
solutions sampled during the Monte Carlo. A part from finite size scaling 
corrections, we observe striking agreement with analytical results, at least 
for the first few root distributions.

%It is useful to rewrite~\eref{d-ensemble} as 
%
%\begin{equation}
%\label{qa-prob}
%\langle{\mathcal O}\rangle=\sum\limits_{\alpha}\rho^{DE}(\alpha)
%\langle\alpha|{\mathcal O}|\alpha\rangle,\quad\textrm{with}\quad 
%\rho^{DE}(\alpha)=\exp(-2\Re{\mathcal E}(\alpha)).
%\end{equation}
%
%Here $\rho^{DE}(\alpha)$ is a diagonal ensemble density matrix, and ${\mathcal E}(\alpha)
%\equiv-\log\langle\alpha|\Psi_0\rangle$, with $\Re$ denoting the real part.

%The N\'eel state is defined as 
%
%\begin{equation}
%\label{neel}
%|N\rangle\equiv\frac{1}{\sqrt{2}}\big(|N_1\rangle
%+|N_2\rangle\big), 
%\end{equation}
%

@@

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.2} discuss the structure of  
its eigenstates (Bethe states) and the associated Bethe equations. 
Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide the exact formulas for the 
local conserved charges of the model.

@@

We stard discussing the overlaps with the N\'eel state. Due to the zero-momentum 
constraint, only parity-invariant Bethe states can have non-zero 
N\'eel overlap~\cite{brockmann-2014,wouters-2014}. The corresponding solutions of the Bethe 
equations~\eref{ba-eq} contain only pairs of rapidities with opposite sign.  Here we denote the 
generic parity-invariant rapidity configuration as $|\{\pm\tilde\lambda_j\}_{j=1}^m,n_\infty
\rangle$, i.e., considering only positive rapidities (as stressed by the tilde in $\tilde
\lambda_j$). Here $m$ is the number of rapidity pairs. Since the N\'eel state is not invariant 
under $SU(2)$ rotations, eigenstates with infinite rapidities can have non-zero N\'eel overlaps.
We denote the number of infinite rapidities as $N_{\infty}$. Note that one has $M=L/2=
N_\infty+2m$. The density of infinite rapidities is denoted as $n_\infty\equiv N_\infty/L$. 
The overlap between the Bethe states and the Neel state $|N\rangle$ reads~\cite{brockmann-2014,
pozsgay-2014} 
%
\begin{equation}
\label{neel-ov}
\frac{\langle N|\{\pm\tilde\lambda_j\}_{j=1}^m,n_\infty\rangle}{|||\{\tilde\lambda_j\}_{j=1}^m,
n_\infty\rangle||}=\frac{\sqrt{2}N_{\infty}!}{\sqrt{(2N_\infty)!}}\left[\prod_{j=1}^m
\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}\right]\sqrt{\frac{\textrm{det}_m(G^+)}{
\textrm{det}_m(G^-)}}.
\end{equation}
%
The matrix $G^\pm$ is  defined as  
%
\begin{equation}
\label{G-pm}
\fl\quad G^{\pm}_{jk}=\delta_{jk}\Big(LK_{1/2}(\tilde\lambda_j)-\sum\limits_{l=1}^mK_1^+(
\tilde\lambda_j,\tilde\lambda_l)\Big)+K_{1}^{\pm}(\tilde\lambda_j,\tilde\lambda_k),
\quad\,j,k=1,\dots,m, 
\end{equation}
%
where 
%
\begin{equation}
\label{K}
K_1^\pm(\lambda,\mu)=K_1(\lambda-\mu)\pm K_1(\lambda+\mu) \quad\textrm{with}\quad 
K_\alpha(\lambda)\equiv\frac{8\alpha}{\lambda^2+4\alpha^2}. 
\end{equation}
%
Note that our definitions of $K_{\alpha}(\lambda)$ differs from the one in 
Ref.~\cite{brockmann-2014}, due to a factor $2$ in the definition of the 
rapidities (cf.~\eref{str-hyp}).

@@

length is $M/2$ beacause of parity invariance. Note also that in determining $q$, strings 

>>>>

\title{The quench action approach in finite integrable spin chains}

\author{Vincenzo Alba and Pasquale Calabrese }
%
\address{$^1$\,SISSA and INFN, via Bonomea 265, 34136 Trieste, Italy. }
%\address{$^2$\,The Rudolf Peierls Centre for Theoretical Physics,
   % Oxford University, Oxford, OX1 3NP, United Kingdom.   }

@@

We consider the problem of constructing the stationary state following a quantum quench, 
using the exact overlaps for finite size integrable models. 
We focus on the isotropic Heisenberg spin chain with initial state N\'eel or Majumdar-Ghosh (dimer),
although the proposed approach is valid for an arbitrary integrable model. 
We consider only those states which do not contain zero-momentum strings because the latter 
are affected by fictitious singularities that are very difficult to take into account. 
We show that the fraction of eigenstates that do not contain zero-momentum strings is vanishing in the thermodynamic limit.
Consequently, restricting to this part of the Hilbert space would lead to vanishing expectation values of local observables. 
It is, however, possible to reconstruct the asymptotic values by properly reweighing the expectations in the considered 
subspace at the price of introducing finite-size corrections. 
We also develop a Monte Carlo sampling of the Hilbert space which allows us to study larger systems. 
We accurately reconstruct the expectation values of the conserved charged and the root distributions in the stationary state, 
which turn out to match the exact thermodynamic results. 
The proposed method can be  implemented even in cases in which an analytic thermodynamic solution is not obtainable.

\end{abstract}

@@

Understanding the out-of-equilibrium dynamics in {\it isolated} quantum 
many-body systems is one of the most intriguing research topics in 
contemporary physics, both experimentally~\cite{bloch-2008,greiner-2002,
kinoshita-2006,hofferberth-2007,trotzky-2012,gring-2012,cheneau-2012,
schneider-2012,kunhert-2013,langen-2013,meinert-2013,fukuhara-2013,
ronzheimer-2013,braun-2014,langen-2015} and theoretically~\cite{polkovnikov-2011,efg-15}. 
The most investigated protocol is that of the {\it quantum quench}, in 
which a system is initially prepared in an eigenstate $|\Psi_0\rangle$ of 
a many-body hamiltonian ${\mathcal H}_0$. Then a global parameter is 
suddenly changed, and the system is let to evolve unitarily under a new 
hamiltonian ${\mathcal H}$. There is now compelling evidence that at long 
times after the quench an equilibrium steady state arises (see for instance 
Ref.~\cite{kinoshita-2006}), although its nature is not fully understood yet. 
In the thermodynamic limit, due to dephasing, it is natural to expect that 
the equilibrium value of any local observable ${\mathcal O}$ is described 
by the so-called diagonal ensemble as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle_{DE}=\sum\limits_{\lambda}|\langle\Psi_0|\lambda
\rangle|^2\langle\lambda|{\mathcal O}|\lambda\rangle,
\end{equation}
%
where the sum is over the eigenstates $|\lambda\rangle$ of the post-quench 
hamiltonian ${\mathcal H}$. Moreover, for generic, i.e., non-integrable, models 
the Eigenstate Thermalization Hypothesis~\cite{deutsch-1991,srednicki-1994} 
(ETH) implies that the diagonal ensemble~\eref{d-ensemble} becomes equivalent 
to the usual Gibbs (thermal) ensemble (a conjecture which has been investigated in many 
numerical studies \cite{tvar,alba-2015a}).

In integrable models, however, the presence of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the 
onset of thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,
rigol-2008} that in this situation the post-quench steady-state can be described 
by the so-called Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved quantities, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
%Note that for the diagonal ensemble~\eref{d-ensemble}  one has trivially that $\langle
%{\mathcal O}\rangle_{DE}=\langle\Psi_0|{\mathcal O}|\Psi_0\rangle$. 
This also provides  useful sum rules that will be of interest in this paper. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2006,cramer-2008,barthel-2008,rossini-2009,calabrese-2011,cazalilla-2012,se-12,mossel-2012a,collura-2013,fagotti-2013,kcc14,kcc14a,sotiriadis-2014,bkc-14}, 
whereas in interacting ones the scenario is 
still not fully settled and some recent works~\cite{ilievski-2015a,cardy-2015} suggest that the $GGE$ 
description of the steady state  is complete provided that the so-called quasi-local 
charges~\cite{prosen-2014,pereira-2014,ilievski-2015,essler-2015} are included in the GGE~\eref{gge}.

Valuable insights into this issue have been provided by the so-called quench-action 
approach~\cite{caux-2013}. For Bethe ansatz integrable models this method allows one 
to construct the diagonal ensemble~\eref{d-ensemble} directly in the thermodynamic 
limit, provided that the overlaps $\langle\lambda|\Psi_0\rangle$ are known. The 
physical idea is that for large system sizes it is possible to approximate the sum 
over the eigenstates in~\eref{d-ensemble} using a saddle point argument. Specifically, 
one starts with rewriting~\eref{d-ensemble} as 
%
\begin{equation}
\label{de1}
\langle{\mathcal O}\rangle_{DE}= \sum\limits_{\lambda}\rho^{DE}\langle\lambda|
{\mathcal O}|\lambda\rangle,\quad\textrm{with}\quad\rho^{DE}=\exp(-2{\rm Re}({\mathcal 
E}(\lambda))), 
\end{equation}
%
where ${\mathcal E}(\lambda)\equiv-\ln\langle\lambda|\Psi_0\rangle$. 
For typical initial states $|\Psi_0\rangle$,
one has ${\mathcal E}\propto L$, with $L$ the system size. This reflects the 
vanishing of the overlaps as $\langle\Psi_0|\lambda\rangle\propto e^{-c L}$.  
As in the standard Thermodynamic Bethe Ansatz (TBA)~\cite{taka-book}, the 
extensivity of ${\mathcal E}$ suggests that in the thermodynamic limit the 
sum in~\eref{de1} is dominated by a saddle point. Remarkably, for some Bethe 
ansatz solvable models and for simple enough initial states, it is possible to 
determine this saddle point {\it analytically}. This has been done successfully, 
for instance, for the quench from the Bose-Einstein condensate
 in the Lieb-Liniger model ~\cite{de-nardis-2014,ga-15,pce-2015,b-15}, 
 for the quench from some product states in the $XXZ$ spin chain \cite{pozsgay-2014A,wouters-2014A}, 
for transport in spin chains \cite{de_luca}, and for some interacting field theories \cite{bse-14}.  
The quench action approach allows in principle also to reconstruct the full relaxation dynamics to the steady state \cite{caux-2013}
as numerically done for the 1D Bose gas~\cite{de-nardis-2015a,de-nardis-2015}.

\paragraph*{Outline of the results.} 
In this paper, by combining exact Bethe ansatz and Monte Carlo techniques, we investigate 
the diagonal ensemble and the quench action approach in {\it finite} size integrable models. 
%
We consider the spin-$1/2$ XXZ spin-chain with $L$ sites, which is defined by the Hamiltonian 
%
\begin{equation}
\label{xxx-ham}
{\mathcal H}\equiv J\sum\limits_{i=1}^L\left[\frac{1}{2}(S_i^+S^-_{i+1} 
+S_i^{-}S_{i+1}^+)+\Delta S_i^zS_{i+1}^z-\frac{1}{4}\right].
\end{equation}
%
Here $S^{\pm}_i\equiv (\sigma_i^x\pm i\sigma_i^y)/2$ are spin operators acting on the 
site $i$, $S_i^z\equiv\sigma_i^z/2$, and $\sigma^{x,y,z}_i$ the Pauli matrices. We 
fix $J=1$ in~\eref{xxx-ham} and use periodic boundary conditions, identifying sites 
$L+1$ and $1$. The total magnetization $S_{T}^z\equiv\sum_iS_i^z=L/2-M$, with $M$ number 
of down spins (particles), commutes with~\eref{xxx-ham}, and it is used to label its 
eigenstates. The $XXZ$ chain is Bethe ansatz solvable~\cite{bethe-1931,kor-book} and its eigenstates 
are in one-to-one correspondence with the solutions of the so-called Bethe equations 
(see~\ref{sec:1.2}). 
%
The non-equilibrium quench dynamics of the XXZ spin chain has been investigated numerically intensively \cite{dmcf-06,bar-09,krs-12,cbp-13,cce-15,fagotti-2014},
the GGE with all the local charges has been explicitly constructed \cite{fagotti-2014,pozsgay-2013,fagotti-2013b}, 
and the quench action solution provided for some initial states \cite{pozsgay-2014A,wouters-2014A}.

For concreteness reasons, we focus on the isotropic Heisenberg point ($XXX$), i.e. Hamiltonian \eref{xxx-ham} with $\Delta=1$, 
but we stress that our approach applies to arbitrary $\Delta$, and  also to arbitrary integrable models. 
We consider the quenches from the zero-momentum N\'eel 
state $|N\rangle$ and the Majumdar-Ghosh state $|MG\rangle$, which are defined as  
%
\begin{equation}
\label{psi0}
|N\rangle\equiv\Big(\frac{|N_1\rangle
+|N_2\rangle}{\sqrt{2}}\Big), \qquad 
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
Here $|N_1\rangle\equiv |\uparrow\downarrow\rangle^{\otimes L/2}$, and $|N_2\rangle
\equiv |\downarrow\uparrow\rangle^{\otimes L/2}$, with $\otimes$ denoting the tensor 
product. Note that both $|N\rangle$ and $|MG\rangle$ are invariant under one site 
translations. Our study relies on the analytical knowledge of the overlaps between 
the N\'eel and Majumdar-Ghosh state and the eigenstates of~\eref{xxx-ham}~\cite{
brockmann-2014,brockmann-2014b,brockmann-2014c,pozsgay-2014a,piroli-2014,mazza-2015}.

We first present a detailed overview of the distribution of the overlaps between 
the initial states and the eigenstates of the model (overlap distribution function). 
Precisely, we provide numerical results for all the overlaps for finite chains up 
to $L\le 38$. Our results are obtained exploiting the analytical formulas 
presented in Ref.~\cite{brockmann-2014} and~\cite{pozsgay-2014a}. Crucially, 
we restrict ourselves to a truncated Hilbert space, considering only eigenstates 
of~\eref{xxx-ham} that do not contain zero-momentum strings. Physically, 
zero-momentum strings correspond to eigenstates amplitudes containing multi-particle  
bound states which are fully delocalized in the chain. From the Bethe ansatz perspective, 
in the thermodynamic limit the presence of zero-momentum strings leads to fictitious 
singularities in the overlap formulas. Dealing with these singularities is 
a formidable task that requires detailed knowledge of the solutions of the 
Bethe equations, and it can be done only in very simple cases, e.g., for small 
chains (as explicitly done for the attractive Lieb-Liniger gas for small number of particles \cite{calabrese-2014}). 
On the other hand, it has been argued that zero-momentum strings are irrelevant for the reconstruction of the 
representative state in the thermodynamic limit~\cite{wouters-2014A}.

For a finite chain, we find that for both the N\'eel and the Majumdar-Ghosh initial states, 
the fraction of eigenstates that do not contain zero-momentum strings 
is vanishing in the thermodynamic limit, meaning that for finite large-enough 
systems the vast majority of the eigenstates contain zero momentum strings. Specifically, 
the total number of eigenstates without zero-momentum strings is given in terms of the 
chain size $L$ by simple combinatorial formulas that we provide. 
%
We investigate the effect of the Hilbert space truncation on the diagonal ensemble focusing on the  
sum rules $\langle Q_j\rangle_{DE}=\langle\Psi_0|Q_j|\Psi_0\rangle$ for the conserved quantities $Q_j$. 
We numerically demonstrate that truncating the diagonal ensemble to 
eigenstates with no zero-momentum strings leads to striking violations of the sum rules,  
as one would naively expect since most of the states with non zero overlaps have not been included. 
Precisely, we numerically observe that the truncated diagonal ensemble 
average $\langle Q_j\rangle$ vanish in the thermodynamic limit reflecting the 
vanishing behavior of the fraction of eigenstates with finite-momentum strings. 
Thus, for finite chains, eigenstates corresponding 
to zero-momentum strings cannot be trivially neglected when considering 
diagonal ensemble averages.

However, this does not imply that the diagonal ensemble cannot be reconstructed  
from the eigenstates without zero-momentum strings and it is not in contrast 
with the previous quench action results in the thermodynamic limit \cite{wouters-2014A,pozsgay-2014A}.
Indeed, we show  that the correct diagonal ensemble 
sum rules can be recovered in the thermodynamic limit by appropriately 
reweighing the contribution of the finite-momentum strings eigenstates.

%The key idea is to sample the eigenstates with no zero-momentum 
%strings with the diagonal ensemble probability measure $\rho^{DE}$ (cf.~\eref{de1}). 
We also develop a Monte Carlo scheme, which is a generalization of 
the approach presented in Ref.~\cite{alba-2015} to simulate the Generalize 
Gibbs Ensemble. The method is based on the knowledge of the N\'eel overlaps 
$\langle\lambda|N\rangle$ and on the knowledge of the Hilbert space structure 
of the $XXX$ chain in the Bethe ansatz formalism. The approach allows us to 
simulate effectively chain with $L\lesssim 60$, although larger systems 
sizes can in principle be reached.

Strikingly, after the reweighing, although for small chains violations of the conserved quantities 
sum rules are present, these violations vanish in the thermodynamic limit 
and the sum rules are restored. This implies that the only effect of the 
Hilbert space truncation is to introduce finite size scaling corrections. 
In the quench action language, this means that the eigenstates corresponding to non 
zero-momentum strings contain enough physical information about saddle point. 
This is numerically confirmed by extracting the so-called saddle point 
root distributions, which in the Bethe ansatz language fully characterize 
the diagonal ensemble averages in the thermodynamic limit. In the numerical  
 approach these are obtained from the histograms of the Bethe ansatz 
solutions sampled during the Monte Carlo. A part from finite size scaling 
corrections, we observe striking agreement with analytical results, at least 
for the first few root distributions.

%It is useful to rewrite~\eref{d-ensemble} as 
%
%\begin{equation}
%\label{qa-prob}
%\langle{\mathcal O}\rangle=\sum\limits_{\alpha}\rho^{DE}(\alpha)
%\langle\alpha|{\mathcal O}|\alpha\rangle,\quad\textrm{with}\quad 
%\rho^{DE}(\alpha)=\exp(-2\Re{\mathcal E}(\alpha)).
%\end{equation}
%
%Here $\rho^{DE}(\alpha)$ is a diagonal ensemble density matrix, and ${\mathcal E}(\alpha)
%\equiv-\ln\langle\alpha|\Psi_0\rangle$, with $\Re$ denoting the real part.

%The N\'eel state is defined as 
%
%\begin{equation}
%\label{N\'eel}
%|N\rangle\equiv\frac{1}{\sqrt{2}}\big(|N_1\rangle
%+|N_2\rangle\big), 
%\end{equation}
%

@@

In this section we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in subsection~\ref{sec:1.2} we discuss the structure of  
its eigenstates (Bethe states) and the associated Bethe equations. 
Subsection~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in subsection~\ref{sec:1.4}. 
Finally, in subsection~\ref{sec:1.5} we provide the exact formulas for the 
local conserved charges of the model.

@@

We start discussing the overlaps with the N\'eel state. 
As shown in Ref. \cite{brockmann-2014,wouters-2014A}, only parity-invariant Bethe states have non-zero overlap with the N\'eel state. 
The corresponding solutions of the Bethe equations~\eref{ba-eq} contain only pairs of rapidities with opposite sign.  
Here we denote the generic parity-invariant rapidity configuration as $|\{\pm\tilde\lambda_j\}_{j=1}^m,n_\infty\rangle$, 
i.e., considering only positive rapidities (as stressed by the tilde in $\tilde\lambda_j$). 
Here $m$ is the number of rapidity pairs. Since the N\'eel state is not invariant 
under $SU(2)$ rotations, eigenstates with infinite rapidities can have non-zero N\'eel overlaps.
We denote the number of infinite rapidities as $N_{\infty}$. Note that one has $M=L/2=
N_\infty+2m$. The density of infinite rapidities is denoted as $n_\infty\equiv N_\infty/L$. 
The overlap between the Bethe states and the N\'eel state $|N\rangle$ reads~\cite{wouters-2014A,
pozsgay-2014a} 
%
\begin{equation}
\label{Neel-ov}
\frac{\langle N|\{\pm\tilde\lambda_j\}_{j=1}^m,n_\infty\rangle}{|||\{\tilde\lambda_j\}_{j=1}^m,
n_\infty\rangle||}=\frac{\sqrt{2}N_{\infty}!}{\sqrt{(2N_\infty)!}}\left[\prod_{j=1}^m
\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}\right]\sqrt{\frac{\textrm{det}_m(G^+)}{
\textrm{det}_m(G^-)}}.
\end{equation}
%
The matrix $G^\pm$ is  defined as  
%
\begin{equation}
\label{G-pm}
\fl\quad G^{\pm}_{jk}=\delta_{jk}\Big(LK_{1/2}(\tilde\lambda_j)-\sum\limits_{l=1}^mK_1^+(
\tilde\lambda_j,\tilde\lambda_l)\Big)+K_{1}^{\pm}(\tilde\lambda_j,\tilde\lambda_k),
\quad\,j,k=1,\dots,m, 
\end{equation}
%
where 
%
\begin{equation}
\label{K}
K_1^\pm(\lambda,\mu)=K_1(\lambda-\mu)\pm K_1(\lambda+\mu) \quad\textrm{with}\quad 
K_\alpha(\lambda)\equiv\frac{8\alpha}{\lambda^2+4\alpha^2}. 
\end{equation}
%
Note that our definitions of $K_{\alpha}(\lambda)$ differs from the one in 
Ref.~\cite{brockmann-2014}, due to a factor $2$ in the definition of the 
rapidities (cf.~\eref{str-hyp}).

@@

length is $M/2$ because of parity invariance. Note also that in determining $q$, strings 
